quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-06-24 paper-quollr.Rmd
Added quollr motivation letter

all_hex_coord
```

@@

umap_hex_id
```

@@

std_df
```

@@

pts_df
```

@@

df_bin_centroids
```

@@

trimesh
```

@@

df_bin
```

@@

```{r, echo=TRUE}
predict_emb(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
```

@@

model_error
```

@@

```{r, echo=TRUE}

df_exe <- comb_data_model(
  highd_data = scurve, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids
  )

df_exe
```

@@

```{r, echo=TRUE}
df_exe <- comb_all_data_model_error(
  highd_data = scurve, 
  nldr_data = scurve_umap, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids, 
  error_data = model_error
  )

df_exe
```

@@

Limb muscle cells of mice in @tabula2018 are examined. There are $1067$ single cells, with $14997$ gene expressions. Following their pre-processing, tSNE was performed using ten principal components. Figure \@ref(fig:limb-rmse) (b) is the reproduction of the published plot. The question is whether this accurately represents the cluster structure in the data. Our method can help here, and also help to provide a better $2-\text{D}$ layout, as needed.

>>>>

head(all_hex_coord, 5)
```

@@

head(umap_hex_id, 5)
```

@@

head(std_df, 5)
```

@@

head(pts_df, 5)
```

@@

head(df_bin_centroids, 5)
```

@@

head(trimesh, 5)
```

@@

head(df_bin, 5)
```

@@

```{r, echo=TRUE}
predict_data <- predict_emb(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )

head(predict_data, 5)
```

@@

head(model_error, 5)
```

@@

```{r, echo=TRUE}
df_exe <- comb_data_model(
  highd_data = scurve, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids
  )
```

@@

```{r, echo=TRUE}
df_exe <- comb_all_data_model_error(
  highd_data = scurve, 
  nldr_data = scurve_umap, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids, 
  error_data = model_error
  )
```

@@

Limb muscle cells of mice in @tabula2018 are examined. There are $1067$ single cells, with $14997$ gene expressions. Following their pre-processing, different NLDR methods were performed using ten principal components. Figure \@ref(fig:limb-rmse) (b) is the reproduction of the published plot. The question is whether this accurately represents the cluster structure in the data. Our method can help here, and also help to provide a better $2-\text{D}$ layout, as needed.

### 2025-06-24 paper-quollr.tex
Added quollr motivation letter

all_hex_coord
\end{verbatim}

\begin{verbatim}
#> # A tibble: 1,440 x 3
#>    hex_poly_id       x       y
#>          <int>   <dbl>   <dbl>
#>  1           1 -0.1    -0.0412
#>  2           1 -0.142  -0.0652
#>  3           1 -0.142  -0.113 
#>  4           1 -0.1    -0.137 
#>  5           1 -0.0584 -0.113 
#>  6           1 -0.0584 -0.0652
#>  7           2 -0.0167 -0.0412
#>  8           2 -0.0584 -0.0652
#>  9           2 -0.0584 -0.113 
#> 10           2 -0.0167 -0.137 
#> # i 1,430 more rows
\end{verbatim}

@@

umap_hex_id
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5,000 x 4
#>     emb1  emb2    ID hexID
#>    <dbl> <dbl> <int> <int>
#>  1 0.707 0.839     1   205
#>  2 0.231 0.401     2   109
#>  3 0.232 0.215     3    65
#>  4 0.790 0.564     4   146
#>  5 0.761 0.551     5   146
#>  6 0.445 0.721     6   172
#>  7 0.900 0.137     7    58
#>  8 0.247 0.392     8   110
#>  9 0.325 0.542     9   141
#> 10 0.278 0.231    10    80
#> # i 4,990 more rows
\end{verbatim}

@@

std_df
\end{verbatim}

\begin{verbatim}
#> # A tibble: 142 x 3
#>    hexID bin_counts std_counts
#>    <int>      <int>      <dbl>
#>  1    21         12     0.16  
#>  2    22         17     0.227 
#>  3    23         22     0.293 
#>  4    24         10     0.133 
#>  5    25          7     0.0933
#>  6    26         12     0.16  
#>  7    27          1     0.0133
#>  8    36         42     0.56  
#>  9    37         42     0.56  
#> 10    38         44     0.587 
#> # i 132 more rows
\end{verbatim}

@@

pts_df
\end{verbatim}

\begin{verbatim}
#> # A tibble: 142 x 2
#>    hexID pts_list  
#>    <int> <list>    
#>  1    21 <int [12]>
#>  2    22 <int [17]>
#>  3    23 <int [22]>
#>  4    24 <int [10]>
#>  5    25 <int [7]> 
#>  6    26 <int [12]>
#>  7    27 <int [1]> 
#>  8    36 <int [42]>
#>  9    37 <int [42]>
#> 10    38 <int [44]>
#> # i 132 more rows
\end{verbatim}

@@

df_bin_centroids
\end{verbatim}

\begin{verbatim}
#> # A tibble: 240 x 5
#>    hexID     c_x     c_y bin_counts std_counts
#>    <int>   <dbl>   <dbl>      <dbl>      <dbl>
#>  1     1 -0.1    -0.0892          0          0
#>  2     2 -0.0167 -0.0892          0          0
#>  3     3  0.0665 -0.0892          0          0
#>  4     4  0.150  -0.0892          0          0
#>  5     5  0.233  -0.0892          0          0
#>  6     6  0.316  -0.0892          0          0
#>  7     7  0.400  -0.0892          0          0
#>  8     8  0.483  -0.0892          0          0
#>  9     9  0.566  -0.0892          0          0
#> 10    10  0.649  -0.0892          0          0
#> # i 230 more rows
\end{verbatim}

@@

trimesh
\end{verbatim}

\begin{verbatim}
#> # A tibble: 653 x 8
#>     from    to  x_from  y_from    x_to    y_to from_count to_count
#>    <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
#>  1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
#>  2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
#>  3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
#>  4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
#>  5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
#>  6    17    33  0.0249 -0.0171  0.0665  0.0550          0        0
#>  7    31    46 -0.1     0.0550 -0.0584  0.127           0        0
#>  8    32    47 -0.0167  0.0550  0.0249  0.127           0        0
#>  9    32    33 -0.0167  0.0550  0.0665  0.0550          0        0
#> 10     4    18  0.150  -0.0892  0.108  -0.0171          0        0
#> # i 643 more rows
\end{verbatim}

@@

df_bin
\end{verbatim}

\begin{verbatim}
#> # A tibble: 142 x 8
#>    hexID      x1    x2    x3         x4        x5       x6        x7
#>    <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
#>  1    21 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105 
#>  2    22 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363
#>  3    23 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153 
#>  4    24 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128 
#>  5    25  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195 
#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
#>  7    27  0.719   1.99 1.70   0.0109    -0.00349  -0.0297  -0.00223 
#>  8    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
#>  9    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
#> 10    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
#> # i 132 more rows
\end{verbatim}

@@

\begin{verbatim}
predict_emb(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5,000 x 4
#>    pred_emb_1 pred_emb_2    ID pred_hb_id
#>         <dbl>      <dbl> <int>      <int>
#>  1      0.691      0.848     1        205
#>  2      0.191      0.416     2        109
#>  3      0.316      0.199     3         66
#>  4      0.774      0.560     4        146
#>  5      0.774      0.560     5        146
#>  6      0.441      0.704     6        172
#>  7      0.941      0.127     7         58
#>  8      0.275      0.416     8        110
#>  9      0.358      0.560     9        141
#> 10      0.316      0.343    10         96
#> # i 4,990 more rows
\end{verbatim}

@@

model_error
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5,000 x 32
#>       ID      x1     x2       x3       x4       x5       x6        x7 pred_hb_id
#>    <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>    <dbl>     <dbl>      <int>
#>  1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923  -0.00128         205
#>  2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362  -0.00535         109
#>  3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962   0.00335          66
#>  4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716   0.00126         146
#>  5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998   0.000677        146
#>  6     6  0.818  0.927  -1.58    -0.00318 -0.00980  0.0989   0.00696         172
#>  7     7  0.910  1.40    1.42     0.00699 -0.0182  -0.0710   0.00966          58
#>  8     8 -0.0691 1.59    0.00239  0.0127  -0.0130   0.0396  -0.000185        110
#>  9     9  0.859  1.59   -0.488   -0.0119   0.00421 -0.00440 -0.00595         141
#> 10    10 -0.727  1.62    0.314    0.00251  0.0177  -0.0755  -0.00369          96
#> # i 4,990 more rows
#> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
#> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
#> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
#> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
#> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>, ...
\end{verbatim}

@@

\begin{verbatim}
df_exe <- comb_data_model(
  highd_data = scurve, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids
  )

df_exe
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5,122 x 8
#>         x1    x2    x3         x4        x5       x6        x7 type 
#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr>
#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model
#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model
#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model
#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model
#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model
#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model
#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model
#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model
#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model
#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model
#> # i 5,112 more rows
\end{verbatim}

@@

\begin{verbatim}
df_exe <- comb_all_data_model_error(
  highd_data = scurve, 
  nldr_data = scurve_umap, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids, 
  error_data = model_error
  )

df_exe
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5,122 x 12
#>         x1    x2    x3         x4        x5       x6        x7 type   emb1  emb2
#>      <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl> <chr> <dbl> <dbl>
#>  1 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105  model    NA    NA
#>  2 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363 model    NA    NA
#>  3 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153  model    NA    NA
#>  4  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172  model    NA    NA
#>  5 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338 model    NA    NA
#>  6 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126 model    NA    NA
#>  7 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780 model    NA    NA
#>  8 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117  model    NA    NA
#>  9 -0.0991  1.70 1.99   0.00103    0.00150   0.00877 -0.000193 model    NA    NA
#> 10  0.295   1.74 1.95  -0.00165    0.000459  0.00330  0.000257 model    NA    NA
#> # i 5,112 more rows
#> # i 2 more variables: sqrt_row_wise_total_error <dbl>, density <dbl>
\end{verbatim}

>>>>

head(all_hex_coord, 5)
\end{verbatim}

\begin{verbatim}
#>   hex_poly_id           x           y
#> 1           1 -0.10000000 -0.04116510
#> 2           1 -0.14163068 -0.06520059
#> 3           1 -0.14163068 -0.11327155
#> 4           1 -0.10000000 -0.13730704
#> 5           1 -0.05836932 -0.11327155
\end{verbatim}

@@

head(umap_hex_id, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 4
#>    emb1  emb2    ID hexID
#>   <dbl> <dbl> <int> <int>
#> 1 0.707 0.839     1   205
#> 2 0.231 0.401     2   109
#> 3 0.232 0.215     3    65
#> 4 0.790 0.564     4   146
#> 5 0.761 0.551     5   146
\end{verbatim}

@@

head(std_df, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 3
#>   hexID bin_counts std_counts
#>   <int>      <int>      <dbl>
#> 1    21         12     0.16  
#> 2    22         17     0.227 
#> 3    23         22     0.293 
#> 4    24         10     0.133 
#> 5    25          7     0.0933
\end{verbatim}

@@

head(pts_df, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 2
#>   hexID pts_list  
#>   <int> <list>    
#> 1    21 <int [12]>
#> 2    22 <int [17]>
#> 3    23 <int [22]>
#> 4    24 <int [10]>
#> 5    25 <int [7]>
\end{verbatim}

@@

head(df_bin_centroids, 5)
\end{verbatim}

\begin{verbatim}
#>   hexID         c_x         c_y bin_counts std_counts
#> 1     1 -0.10000000 -0.08923607          0          0
#> 2     2 -0.01673864 -0.08923607          0          0
#> 3     3  0.06652271 -0.08923607          0          0
#> 4     4  0.14978407 -0.08923607          0          0
#> 5     5  0.23304543 -0.08923607          0          0
\end{verbatim}

@@

head(trimesh, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 8
#>    from    to  x_from  y_from    x_to    y_to from_count to_count
#>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
#> 1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
#> 2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
#> 3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
#> 4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
#> 5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
\end{verbatim}

@@

head(df_bin, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 8
#>   hexID      x1    x2    x3         x4        x5       x6        x7
#>   <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
#> 1    21 -0.992   1.91  1.11 -0.000427   0.000624  0.00749  0.00105 
#> 2    22 -0.906   1.93  1.41 -0.0000183  0.00331  -0.0204  -0.000363
#> 3    23 -0.680   1.93  1.72 -0.000810  -0.00259  -0.00449  0.00153 
#> 4    24 -0.272   1.93  1.96  0.00251    0.00668  -0.0460   0.00128 
#> 5    25  0.0760  1.93  2.00  0.00876    0.00447   0.00851 -0.00195
\end{verbatim}

@@

\begin{verbatim}
predict_data <- predict_emb(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )

head(predict_data, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 4
#>   pred_emb_1 pred_emb_2    ID pred_hb_id
#>        <dbl>      <dbl> <int>      <int>
#> 1      0.691      0.848     1        205
#> 2      0.191      0.416     2        109
#> 3      0.316      0.199     3         66
#> 4      0.774      0.560     4        146
#> 5      0.774      0.560     5        146
\end{verbatim}

@@

head(model_error, 5)
\end{verbatim}

\begin{verbatim}
#> # A tibble: 5 x 32
#>      ID      x1     x2       x3       x4       x5      x6        x7 pred_hb_id
#>   <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>   <dbl>     <dbl>      <int>
#> 1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923 -0.00128         205
#> 2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362 -0.00535         109
#> 3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962  0.00335          66
#> 4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716  0.00126         146
#> 5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998  0.000677        146
#> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
#> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
#> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
#> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
#> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
#> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
#> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
\end{verbatim}

@@

\begin{verbatim}
df_exe <- comb_data_model(
  highd_data = scurve, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids
  )
\end{verbatim}

@@

\begin{verbatim}
df_exe <- comb_all_data_model_error(
  highd_data = scurve, 
  nldr_data = scurve_umap, 
  model_highd = df_bin, 
  model_2d = df_bin_centroids, 
  error_data = model_error
  )
\end{verbatim}
